---
title: Scripts for the "Geographies and Changes in US Homeownership from 1970 to 2020" paper
author: "Kanglin (Connie) Chen, PhD"
date: "2025-02-16"
output: html_document
---

# Figure 1. Owner-occupied Housing Units (N and %) at the US National Level, 1970-2020
## Note: We have typed national-level housing data in the scripts so that we don't need to import data at this step
```{r}
# Load necessary libraries
library(ggplot2)
library(scales)  # For percentage formatting and number formatting

# Create a data frame from your provided data
medians_df <- data.frame(
  Year = c(1970, 1980, 1990, 2000, 2010, 2020),
  Median_N = c(39811200, 51794545, 59024811, 69815753, 75986074, 80051358),
  Median_Percent = c(0.580, 0.644, 0.642, 0.662, 0.651, 0.631)  # Converted to proportions
)

# Set axis ranges
left_y_range <- c(35000000, 120000000)  # Range for Median_N
right_y_range <- c(0.55, 0.68)   # Range for Median_Percent in proportion

# Calculate scaling factor to align the ranges
scaling_factor <- diff(left_y_range) / diff(right_y_range)

# Scale the percentage data for plotting
medians_df$Scaled_Percent <- (medians_df$Median_Percent - min(right_y_range)) * scaling_factor + min(left_y_range)

# Create the dual-axis line chart
ggplot(medians_df, aes(x = Year)) +
  # First line (Median N)
  geom_line(aes(y = Median_N), color = "#2E7265", size = 1.2) +
  geom_point(aes(y = Median_N), color = "#2E7265", size = 4) +
  
  # Add text labels for Median N
  geom_text(aes(y = Median_N, label = paste0(round(Median_N / 1e6, 1), "M")), 
            vjust = -1, color = "#2E7265", size = 4) +
  
  # Second line (Median Percent, scaled)
  geom_line(aes(y = Scaled_Percent), color = "#DF4649", linetype = "dashed", size = 1.2) +
  geom_point(aes(y = Scaled_Percent), color = "#DF4649", size = 4, shape = 17) +
  
  # Add text labels for Median Percent (scaled back to original %)
  geom_text(aes(y = Scaled_Percent, 
                label = paste0(round(Median_Percent * 100, 1), "%")), 
            vjust = -1, color = "#DF4649", size = 4) +
  
  # Customize scales and add secondary y-axis
  scale_y_continuous(
    name = "Owner-occupied Housing Units (N)", 
    limits = left_y_range,  # Set range for left y-axis
    labels = label_number(scale = 1e-6, suffix = "M"),  # Format as millions (e.g., 40M, 60M)
    sec.axis = sec_axis(
      ~ (. - min(left_y_range)) / scaling_factor + min(right_y_range), 
      name = "Homeownership Rate (%)",
      labels = percent_format(accuracy = 0.1)  # Format as percentage
    )
  ) +
  
  # Labels and theme
  labs(
    title = "Owner-occupied Housing Units, N & %",
    x = "Year",
    y = "Owner-occupied Housing Units (N)"
  ) +
  theme_minimal() +
  theme(
    axis.title.y.left = element_text(size = 14, color = "#2E7265"),  # Left y-axis label color
    axis.title.y.right = element_text(size = 14, color = "#DF4649"), # Right y-axis label color
    axis.text.y.left = element_text(size = 12, color = "#2E7265"),   # Left y-axis tick color
    axis.text.y.right = element_text(size = 12, color = "#DF4649"),  # Right y-axis tick color
    axis.title.x = element_text(size = 14),
    axis.text.x = element_text(size = 14, color="black"),
    plot.title = element_text(size = 16, hjust = 0.5)
  )
```

# Figure 2. Correlations between the Number of Owner-occupied Housing Units (N) and Homeownership Rate (%) among the US Counties in 1970 and 2020
## Set the file path
```{r}
file_path <- "...\\Homeownership_1970_2020_County_Level_Data.csv" # Convert the provided "Homeownership_1970_2020_County_Level_Data.xlsx" file to .csv file. Update with correct path 
Homeownership <- read.csv(file_path, header = TRUE, stringsAsFactors = FALSE)
head(Homeownership)
attach(Homeownership)
```

## Figure 2: 1970 vs 2020: Owner-occupied Housing Units N
```{r}
# Load the ggplot2 library
library(ggplot2)

# Calculate the correlation coefficient
correlation <- cor(Homeownership$N_1970, Homeownership$N_2020, use = "complete.obs")

# Define color classes and breaks based on the specified ranges
Homeownership$Chg_Rate_70_20_class <- cut(
  Homeownership$Chg_Rate_70_20,
  breaks = c(-Inf, -0.05, -0.001, 0.001, 0.05, Inf),
  labels = c("≤ -5.0%", "-4.9%--0.1%", "No Statistical Change", "0.2%-5.0%", "≥ 5.1%"),
  include.lowest = TRUE
)

# Define the colors corresponding to the classes
colors <- c(
  "≤ -5.0%" ="#DF4649", 
  "-4.9%--0.1%" = "#E87D7F",        
  "No Statistical Change" = "#D6D2CE",         
  "0.2%-5.0%" = "#6D9F95",           
  "≥ 5.1%" = "#2E7265"
)

# Create the scatter plot
ggplot(data = Homeownership, aes(x = N_1970, y = N_2020, color = Chg_Rate_70_20_class)) +
  geom_point(size = 2, alpha = 0.9) + # Adjust transparency with alpha (0.5 makes points semi-transparent)
  scale_color_manual(values = colors) + # Apply the custom color scale
  labs(
    x = "Owner-occupied Housing Units (N), 1970",  # X-axis label
    y = "Owner-occupied Housing Units (N), 2020",  # Y-axis label
    subtitle = paste("Correlation: ", round(correlation, 2)),  # Display the correlation
    color = "Differences in \nOwner-occupied \nHousing Units (%), \n1970-2020"  # Legend title
  ) +
  theme_minimal() + # Clean theme
  theme(
    axis.title.x = element_text(size = 14), # Customize x-axis legend
    axis.title.y = element_text(size = 14), # Customize y-axis legend
    axis.text.x = element_text(size = 11),  # Increase font size for x-axis tick labels
    axis.text.y = element_text(size = 11),  # Increase font size for y-axis tick labels
    legend.title = element_text(size = 13), # Legend title size
    legend.text = element_text(size = 11),  # Legend text size
    plot.subtitle = element_text(size = 14, hjust = 0.5) # Center the subtitle
  ) +
  geom_smooth(method = "lm", color = "grey", linetype = "dashed", se = FALSE) # Add trend line
```

## Figure 2: 1970 vs 2020: Owner-occupied Housing Units %
```{r}
# Load necessary libraries
library(ggplot2)
library(scales) # For percent formatting

# Calculate the correlation coefficient
correlation <- cor(Homeownership$Rate_1970, Homeownership$Rate_2020, use = "complete.obs")

# Define color classes and breaks based on the specified ranges
Homeownership$Chg_N_70_20_class <- cut(
  Homeownership$Chg_N_70_20,
  breaks = c(-126304, 367, 1775, 4223, 12107, 835234),
  labels = c("Q1: -126304-367", "Q2: 368-1775", "Q3: 1776-4223", "Q4: 4224-12107", "Q5: 12108-835234"),
  include.lowest = TRUE
)

# Define the colors corresponding to the classes
colors <- c(
  "Q1: -126304-367" = "#E66101",  
  "Q2: 368-1775" = "#FDB863",         
  "Q3: 1776-4223" = "#FFEBBE",           
  "Q4: 4224-12107" = "#B2ABD2",            
  "Q5: 12108-835234" =  "#5E3C99"     
)

# Create the scatter plot
ggplot(data = Homeownership, aes(x = Rate_1970, y = Rate_2020, color = Chg_N_70_20_class)) +
  geom_point(size = 2, alpha = 0.9) + # Adjust transparency with alpha
  scale_color_manual(values = colors) + # Apply the custom color scale
  scale_x_continuous(labels = percent_format(accuracy = 0.1)) + # Convert x-axis labels to percentage
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) + # Convert y-axis labels to percentage
  labs(
    x = "Owner-occupied Housing Units (%), 1970",  # X-axis label
    y = "Owner-occupied Housing Units (%), 2020",  # Y-axis label
    subtitle = paste("Correlation: ", round(correlation, 2)),  # Display the correlation
    color = "Differences in \nOwner-occupied \nHousing Units (N), \n1970-2020"  # Legend title
  ) +
  theme_minimal() + # Clean theme
  theme(
    axis.title.x = element_text(size = 14), # Customize x-axis label
    axis.title.y = element_text(size = 14), # Customize y-axis label
    axis.text.x = element_text(size = 11),  # Increase font size for x-axis tick labels
    axis.text.y = element_text(size = 11),  # Increase font size for y-axis tick labels
    legend.title = element_text(size = 13), # Legend title size
    legend.text = element_text(size = 11),  # Legend text size
    plot.subtitle = element_text(size = 14, hjust = 0.5) # Center the subtitle
  ) +
  geom_smooth(method = "lm", color = "grey", linetype = "dashed", se = FALSE) # Add trend line
```

## Figure 2: % vs N: Owner-occupied Housing Units in 1970
```{r}
# Load necessary libraries
library(ggplot2)
library(scales) # For percent formatting

# Calculate the correlation coefficient
correlation <- cor(Homeownership$Rate_1970, Homeownership$N_1970, use = "complete.obs")

# Create the scatter plot
ggplot(data = Homeownership, aes(x = Rate_1970, y = N_1970)) +
  geom_point(size = 2, alpha = 0.9, color = "#893238E6") + # Set all points to the same color
  scale_x_continuous(
    labels = percent_format(accuracy = 0.1),               # Convert x-axis labels to percentage with one decimal
    expand = expansion(mult = c(0.02, 0.02))               # Add space at edges to avoid clipping
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.02, 0.02))               # Add space at edges to avoid clipping
  ) +
  labs(
    x = "Owner-occupied Housing Units (%), 1970",          # X-axis label
    y = "Owner-occupied Housing Units (N), 1970",          # Y-axis label
    subtitle = paste("Correlation: ", round(correlation, 2)) # Display the correlation
  ) +
  theme_minimal() + # Clean theme
  theme(
    axis.title.x = element_text(size = 14),        # Customize x-axis label size
    axis.title.y = element_text(size = 14),        # Customize y-axis label size
    axis.text.x = element_text(size = 11),         # Increase font size for x-axis tick labels
    axis.text.y = element_text(size = 11),         # Increase font size for y-axis tick labels
    plot.subtitle = element_text(size = 14, hjust = 0.5), # Center-align subtitle
    legend.position = "none",                      # Remove legend
    plot.margin = unit(c(1, 1, 1, 1), "cm")        # Adjust plot margins (top, right, bottom, left)
  ) +
  geom_smooth(
    method = "lm", color = "lightgrey", linetype = "dashed", se = FALSE # Add trend line
  )
```

## Figure 2: % vs N: Owner-occupied Housing Units in 2020
```{r}
# Load necessary libraries
library(ggplot2)
library(scales) # For percent formatting

# Calculate the correlation coefficient
correlation <- cor(Homeownership$Rate_2020, Homeownership$N_2020, use = "complete.obs")

# Create the scatter plot
ggplot(data = Homeownership, aes(x = Rate_2020, y = N_2020)) +
  geom_point(size = 2, alpha = 0.9, color = "#748C5AE6") + # Set all points to the same color
  scale_x_continuous(
    labels = percent_format(accuracy = 0.1),               # Convert x-axis labels to percentage with one decimal
    expand = expansion(mult = c(0.02, 0.02))               # Add space at edges to avoid clipping
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.02, 0.02))               # Add space at edges to avoid clipping
  ) +
  labs(
    x = "Owner-occupied Housing Units (%), 2020",          # X-axis label
    y = "Owner-occupied Housing Units (N), 2020",          # Y-axis label
    subtitle = paste("Correlation: ", round(correlation, 2)) # Display the correlation
  ) +
  theme_minimal() + # Clean theme
  theme(
    axis.title.x = element_text(size = 14),        # Customize x-axis label size
    axis.title.y = element_text(size = 14),        # Customize y-axis label size
    axis.text.x = element_text(size = 11),         # Increase font size for x-axis tick labels
    axis.text.y = element_text(size = 11),         # Increase font size for y-axis tick labels
    plot.subtitle = element_text(size = 14, hjust = 0.5), # Center-align subtitle
    legend.position = "none",                      # Remove legend
    plot.margin = unit(c(1, 1, 1, 1), "cm")        # Adjust plot margins (top, right, bottom, left)
  ) +
  geom_smooth(
    method = "lm", color = "lightgrey", linetype = "dashed", se = FALSE # Add trend line
  )
```

# Figure 3. Data Distribution of Owner-occupied Housing Units (N and %) Among US Counties, 1970–2020 (Total number of counties=3,115)
## Figure 3: Data Distribution of Owner-occupied Housing Units (N) Among US Counties, 1970–2020
```{r}
# Load necessary libraries
library(ggplot2)
library(reshape2)
library(dplyr)

# Select the columns to include in the boxplot
columns <- c("N_1970", "N_1980", "N_1990", "N_2000", "N_2010", "N_2020")
boxplot_data <- Homeownership[, columns]

# Melt the data into long format for ggplot2
long_data <- melt(boxplot_data, variable.name = "Year", value.name = "Value")

# Rename the x-axis labels
long_data$Year <- factor(long_data$Year, 
                         levels = columns,
                         labels = c("1970", "1980", "1990", "2000", "2010", "2020"))

# Calculate the whiskers (Q1 - 1.5 × IQR and Q3 + 1.5 × IQR)
whisker_positions <- long_data %>%
  group_by(Year) %>%
  summarise(
    Q1 = quantile(Value, 0.25, na.rm = TRUE),
    Q3 = quantile(Value, 0.75, na.rm = TRUE),
    Min = min(Value, na.rm = TRUE),  # Actual minimum
    Max = max(Value, na.rm = TRUE),  # Actual maximum
    lower = pmax(Q1 - 1.5 * (Q3 - Q1), min(Value, na.rm = TRUE)),  # Adjusted lower whisker
    upper = Q3 + 1.5 * (Q3 - Q1)  # Upper whisker
  ) %>%
  ungroup()

# Define colors for the boxplots with 80% transparency
colors <- c("#893238E6", "#EE7422E6", "#D9C264E6", "#8EA9A4E6", "#557F99E6", "#748C5AE6")

# Determine y-axis range based on whisker bounds
y_axis_range <- range(whisker_positions$lower, whisker_positions$upper, na.rm = TRUE)

# Create the boxplot with adjustments
ggplot(long_data, aes(x = Year, y = Value, fill = Year)) +
  geom_segment(data = whisker_positions, aes(x = as.numeric(Year) - 0.3, 
                                             xend = as.numeric(Year) + 0.3, 
                                             y = lower, 
                                             yend = lower), 
               color = "black", linewidth = 0.7) +  # Add horizontal line at lower whisker
  geom_segment(data = whisker_positions, aes(x = as.numeric(Year) - 0.3, 
                                             xend = as.numeric(Year) + 0.3, 
                                             y = upper, 
                                             yend = upper), 
               color = "black", linewidth = 0.7) +  # Add horizontal line at upper whisker
  geom_segment(data = whisker_positions, aes(x = as.numeric(Year), 
                                             xend = as.numeric(Year), 
                                             y = lower, 
                                             yend = upper), 
               color = "black", linewidth = 0.7) +  # Add vertical whiskers
  geom_boxplot(outlier.shape = NA, coef = 1.5) +    # Add boxes on top of whiskers to cover them
  scale_fill_manual(values = colors) +              # Apply custom colors with transparency
  scale_y_continuous(limits = y_axis_range) +       # Adjust y-axis to whisker range
  scale_x_discrete() +                              # Correct x-axis as discrete
  labs(
    x = "Year",  # X-axis label
    y = "Owner-occupied Housing Units (N)",
    title = "Homeownership (N): 1970-2020"
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14),       # Customize x-axis title font size
    axis.title.y = element_text(size = 14),       # Customize y-axis title font size
    axis.text.x = element_text(size = 14, angle = 0, hjust = 0.5, color = "black"), # Customize x-axis tick labels
    axis.text.y = element_text(size = 11),        # Customize y-axis tick labels
    legend.position = "none",                     # Remove legend since fill is self-explanatory
    panel.grid.major.x = element_blank(),         # Remove major vertical grid lines
    panel.grid.minor.x = element_blank()          # Remove minor vertical grid lines
  )
```

##Figure 3: Data Distribution of Owner-occupied Housing Units (%) Among US Counties, 1970–2020
```{r}
# Load necessary libraries
library(ggplot2)
library(reshape2)
library(dplyr)
library(scales) # For percent formatting

# Select the columns to include in the boxplot
columns <- c("Rate_1970", "Rate_1980", "Rate_1990", "Rate_2000", "Rate_2010", "Rate_2020")
boxplot_data <- Homeownership[, columns]

# Melt the data into long format for ggplot2
long_data <- melt(boxplot_data, variable.name = "Year", value.name = "Value")

# Rename the x-axis labels
long_data$Year <- factor(long_data$Year, 
                         levels = columns,
                         labels = c("1970", "1980", "1990", "2000", "2010", "2020"))

# Calculate the whiskers (Q1 - 1.5 × IQR and Q3 + 1.5 × IQR)
whisker_positions <- long_data %>%
  group_by(Year) %>%
  summarise(
    Q1 = quantile(Value, 0.25, na.rm = TRUE),
    Q3 = quantile(Value, 0.75, na.rm = TRUE),
    Min = min(Value, na.rm = TRUE),  # Actual minimum
    Max = max(Value, na.rm = TRUE),  # Actual maximum
    lower = pmax(Q1 - 1.5 * (Q3 - Q1), min(Value, na.rm = TRUE)),  # Adjusted lower whisker
    upper = Q3 + 1.5 * (Q3 - Q1)  # Upper whisker
  ) %>%
  ungroup()

# Define colors for the boxplots with 80% transparency
colors <- c("#893238E6", "#EE7422E6", "#D9C264E6", "#8EA9A4E6", "#557F99E6", "#748C5AE6")

# Determine y-axis range based on whisker bounds
y_axis_range <- range(whisker_positions$lower, whisker_positions$upper, na.rm = TRUE)

# Create the boxplot with adjustments
ggplot(long_data, aes(x = Year, y = Value, fill = Year)) +
  geom_segment(data = whisker_positions, aes(x = as.numeric(Year) - 0.3, 
                                             xend = as.numeric(Year) + 0.3, 
                                             y = lower, 
                                             yend = lower), 
               color = "black", linewidth = 0.7) +  # Add horizontal line at lower whisker
  geom_segment(data = whisker_positions, aes(x = as.numeric(Year) - 0.3, 
                                             xend = as.numeric(Year) + 0.3, 
                                             y = upper, 
                                             yend = upper), 
               color = "black", linewidth = 0.7) +  # Add horizontal line at upper whisker
  geom_segment(data = whisker_positions, aes(x = as.numeric(Year), 
                                             xend = as.numeric(Year), 
                                             y = lower, 
                                             yend = upper), 
               color = "black", linewidth = 0.7) +  # Add vertical whiskers
  geom_boxplot(outlier.shape = NA, coef = 1.5) + # Add boxes on top of whiskers to cover them
  scale_fill_manual(values = colors) +           # Apply custom colors with transparency
  scale_y_continuous(
    limits = y_axis_range,                       # Adjust y-axis to whisker range
    labels = percent_format(accuracy = 0.1)      # Format y-axis as percentage with one decimal
  ) +
  scale_x_discrete() +                           # Correct x-axis as discrete
  labs(
    x = "Year",  # X-axis label
    y = "Owner-occupied Housing Units (%)",      # Update y-axis label to reflect percentage
    title = "Percentage of Owner-occupied Housing Units from 1970 to 2020"
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14),       # Customize x-axis title font size
    axis.title.y = element_text(size = 14),       # Customize y-axis title font size
    axis.text.x = element_text(size = 14, angle = 0, hjust = 0.5, color = "black"), # Customize x-axis tick labels
    axis.text.y = element_text(size = 11),        # Customize y-axis tick labels
    legend.position = "none",                     # Remove legend since fill is self-explanatory
    panel.grid.major.x = element_blank(),         # Remove major vertical grid lines
    panel.grid.minor.x = element_blank()          # Remove minor vertical grid lines
  )
```

## Create the table of the number of outliers for Figure 3
```{r}
# Load necessary library
library(dplyr)

# Select the columns for both Rates and Numbers
Rate_columns <- c("Rate_1970", "Rate_1980", "Rate_1990", "Rate_2000", "Rate_2010", "Rate_2020")
N_columns <- c("N_1970", "N_1980", "N_1990", "N_2000", "N_2010", "N_2020")

# Combine the selected columns
all_columns <- c(Rate_columns, N_columns)
boxplot_data <- Homeownership[, all_columns]

# Function to calculate the number of outliers for a column
calculate_outliers <- function(data_column) {
  Q1 <- quantile(data_column, 0.25, na.rm = TRUE)  # First quartile
  Q3 <- quantile(data_column, 0.75, na.rm = TRUE)  # Third quartile
  IQR <- Q3 - Q1  # Interquartile range
  
  # Define lower and upper bounds
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  
  # Count lower and upper outliers
  lower_outliers <- sum(data_column < lower_bound, na.rm = TRUE)
  upper_outliers <- sum(data_column > upper_bound, na.rm = TRUE)
  
  return(c(lower_outliers = lower_outliers, upper_outliers = upper_outliers))
}

# Apply the function to each column
outliers_summary <- sapply(boxplot_data, calculate_outliers)

# Convert results to a data frame for readability
outliers_summary_df <- as.data.frame(t(outliers_summary))
colnames(outliers_summary_df) <- c("Lower_Outliers", "Upper_Outliers")

# Add a new column to indicate whether the data is Percent or N
outliers_summary_df$Category <- rep(c("Rate", "Number"), each = length(Rate_columns)) 

# Print the results
print(outliers_summary_df)
```

# Figure 4. Within-state variability of owner-occupied Housing Units (N and %) at the US County Level, 1970 and 2020
## Figure 4: Owner-occupied housing units (N), 1970
```{r}
# Plot for 1970_N using the median of N of counties
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readr)
library(scales)  # For formatting large numbers

# Load the dataset
file_path <- "...\\Homeownership_1970_2020_County_Level_Data.csv"  # Convert the provided "Homeownership_1970_2020_County_Level_Data.xlsx" file to .csv file. Update with correct path
data <- read_csv(file_path)

# Convert STATE to a factor
data$STATE <- as.factor(data$STATE)

# Count the number of counties per state and calculate the median for sorting
state_summary <- data %>%
  group_by(STATE) %>%
  summarise(County_Count = n(),
            Median_1970_N = median(`1970_N`, na.rm = TRUE))  # Compute median for sorting

# Create a new x-axis label with state name and county count
state_summary <- state_summary %>%
  mutate(STATE_label = paste0(STATE, " (", County_Count, ")"))

# Reorder the factor levels for STATE_label based on the median 1970_N (highest to lowest)
state_summary <- state_summary %>%
  arrange(desc(Median_1970_N))  # Sort by median 1970_N

# Ensure STATE_label follows this sorted order
state_summary$STATE_label <- factor(state_summary$STATE_label, levels = state_summary$STATE_label)

# Merge back to the main data for plotting
data <- left_join(data, state_summary, by = "STATE")

# Ensure STATE_label is a factor with correct ordering
data$STATE_label <- factor(data$STATE_label, levels = state_summary$STATE_label)

# Create the box plot (Ordered by Median 1970_N per State)
ggplot(data, aes(x = STATE_label, y = `1970_N`, fill = as.factor(STATE))) +  
  geom_boxplot(outlier.shape = NA) +  # Remove outliers from box plot
  labs(title = "Owner-occupied Housing Units (N), 1970",
       x = "State",
       y = "N") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12),  # Rotate x-axis labels
        axis.text.y = element_text(size = 14),  
        axis.title.x = element_text(size = 16, face = "bold"),  
        axis.title.y = element_text(size = 16, face = "bold"),  
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5)) +  
  scale_fill_discrete() +  # Default ggplot2 colors
  guides(fill = "none") +  # Remove legend
  scale_y_continuous(limits = c(0, 200000), breaks = seq(0, 200000, by = 100000), labels = comma)  # Format y-axis
```

## Figure 4: Owner-occupied housing units (N), 2020
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readr)
library(scales)  # For formatting large numbers

# Load the dataset
file_path <- "...\\Homeownership_1970_2020_County_Level_Data.csv"  # Convert the provided "Homeownership_1970_2020_County_Level_Data.xlsx" file to .csv file. Update with correct path
data <- read_csv(file_path)

# Convert STATE to a factor
data$STATE <- as.factor(data$STATE)

# Count the number of counties per state and calculate the median for sorting
state_summary <- data %>%
  group_by(STATE) %>%
  summarise(County_Count = n(),
            Median_2020_N = median(`2020_N`, na.rm = TRUE))  # Compute median for sorting

# Create a new x-axis label with state name and county count
state_summary <- state_summary %>%
  mutate(STATE_label = paste0(STATE, " (", County_Count, ")"))

# Reorder the factor levels for STATE_label based on the median 2020_N (highest to lowest)
state_summary <- state_summary %>%
  arrange(desc(Median_2020_N))  # Sort by median 2020_N

# Ensure STATE_label follows this sorted order
state_summary$STATE_label <- factor(state_summary$STATE_label, levels = state_summary$STATE_label)

# Merge back to the main data for plotting
data <- left_join(data, state_summary, by = "STATE")

# Ensure STATE_label is a factor with correct ordering
data$STATE_label <- factor(data$STATE_label, levels = state_summary$STATE_label)

# Create the box plot (Ordered by Median 2020_N per State)
ggplot(data, aes(x = STATE_label, y = `2020_N`, fill = as.factor(STATE))) +  
  geom_boxplot(outlier.shape = NA) +  # Remove outliers from box plot
  labs(title = "Owner-occupied Housing Units (N), 2020",
       x = "State",
       y = "N (2020)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12),  # Rotate x-axis labels
        axis.text.y = element_text(size = 14),  
        axis.title.x = element_text(size = 16, face = "bold"),  
        axis.title.y = element_text(size = 16, face = "bold"),  
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5)) +  
  scale_fill_discrete() +  # Default ggplot2 colors
  guides(fill = "none") +  # Remove legend
  scale_y_continuous(limits = c(0, 200000), breaks = seq(0, 200000, by = 100000), labels = comma)  # Format y-axis
```

## Figure 4: Owner-occupied housing units (%), 1970
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readr)
library(scales)  # For formatting percentages

# Load the dataset
file_path <- "...\\Homeownership_1970_2020_County_Level_Data.csv"  # Convert the provided "Homeownership_1970_2020_County_Level_Data.xlsx" file to .csv file. Update with correct path
data <- read_csv(file_path)

# Convert STATE to a factor
data$STATE <- as.factor(data$STATE)

# Count the number of counties per state and calculate the median for sorting
state_summary <- data %>%
  group_by(STATE) %>%
  summarise(County_Count = n(),
            Median_1970_Percent = median(`1970_Percent`, na.rm = TRUE))  # Compute median for sorting

# Create a new x-axis label with state name and county count
state_summary <- state_summary %>%
  mutate(STATE_label = paste0(STATE, " (", County_Count, ")"))

# Reorder the factor levels for STATE_label based on the median 1970_Percent (highest to lowest)
state_summary <- state_summary %>%
  arrange(desc(Median_1970_Percent))  # Sort by median 1970_Percent

# Ensure STATE_label follows this sorted order
state_summary$STATE_label <- factor(state_summary$STATE_label, levels = state_summary$STATE_label)

# Merge back to the main data for plotting
data <- left_join(data, state_summary, by = "STATE")

# Ensure STATE_label is a factor with correct ordering
data$STATE_label <- factor(data$STATE_label, levels = state_summary$STATE_label)

# Create the box plot (Ordered by Median 1970_Percent per State)
ggplot(data, aes(x = STATE_label, y = `1970_Percent`, fill = as.factor(STATE))) +  
  geom_boxplot(outlier.shape = NA) +  # Remove outliers from box plot
  labs(title = "Owner-occupied Housing Units (%), 1970",
       x = "State",
       y = "%") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12),  # Rotate x-axis labels
        axis.text.y = element_text(size = 14),  
        axis.title.x = element_text(size = 16, face = "bold"),  
        axis.title.y = element_text(size = 16, face = "bold"),  
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5)) +  
  scale_fill_discrete() +  # Default ggplot2 colors
  guides(fill = "none") +  # Remove legend
  scale_y_continuous(limits = c(0.3, 0.9), breaks = seq(0, 1, by = 0.2), labels = percent)  # Format y-axis as percentages
```

## Figure 4: Owner-occupied housing units (%), 2020
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readr)
library(scales)  # For formatting percentages

# Load the dataset
file_path <- "...\\Homeownership_1970_2020_County_Level_Data.csv"  # Convert the provided "Homeownership_1970_2020_County_Level_Data.xlsx" file to .csv file. Update with correct path
data <- read_csv(file_path)

# Convert STATE to a factor
data$STATE <- as.factor(data$STATE)

# Count the number of counties per state and calculate the median for sorting
state_summary <- data %>%
  group_by(STATE) %>%
  summarise(County_Count = n(),
            Median_2020_Percent = median(`2020_Percent`, na.rm = TRUE))  # Compute median for sorting

# Create a new x-axis label with state name and county count
state_summary <- state_summary %>%
  mutate(STATE_label = paste0(STATE, " (", County_Count, ")"))

# Reorder the factor levels for STATE_label based on the median 2020_Percent (highest to lowest)
state_summary <- state_summary %>%
  arrange(desc(Median_2020_Percent))  # Sort by median 2020_Percent

# Ensure STATE_label follows this sorted order
state_summary$STATE_label <- factor(state_summary$STATE_label, levels = state_summary$STATE_label)

# Merge back to the main data for plotting
data <- left_join(data, state_summary, by = "STATE")

# Ensure STATE_label is a factor with correct ordering
data$STATE_label <- factor(data$STATE_label, levels = state_summary$STATE_label)

# Create the box plot (Ordered by Median 2020_Percent per State)
ggplot(data, aes(x = STATE_label, y = `2020_Percent`, fill = as.factor(STATE))) +  
  geom_boxplot(outlier.shape = NA) +  # Remove outliers from box plot
  labs(title = "Owner-occupied Housing Units (%), 2020",
       x = "State",
       y = "%") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12),  # Rotate x-axis labels
        axis.text.y = element_text(size = 14),  
        axis.title.x = element_text(size = 16, face = "bold"),  
        axis.title.y = element_text(size = 16, face = "bold"),  
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5)) +  
  scale_fill_discrete() +  # Default ggplot2 colors
  guides(fill = "none") +  # Remove legend
  scale_y_continuous(limits = c(0.3, 0.9), breaks = seq(0, 1, by = 0.2), labels = percent)  # Format y-axis as percentages
```

